---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Unzip and read the activity data, convert the "date" field to POSIX date type.
```{r}
data <- read.csv(unz("activity.zip", 'activity.csv'))
data$date <- as.POSIXct(data$date)
str(data)
```

## What is mean total number of steps taken per day?

* filter out the rows with missing values
* calculate the total number of steps per day
* plot a histogram of steps per day 

```{r}
stepsReported <- data[!is.na(data$steps),]
stepsByDate <- aggregate(stepsReported$steps, list(date = stepsReported$date), sum)
hist(stepsByDate$x, xlab = "steps", main = "Total number of steps taken each day")
```

* calculate mean of total steps per day
```{r}
mean(stepsByDate$x)
```

* calculate median of total steps per day
```{r}
median(stepsByDate$x)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
