---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

* unzip and read the activity data
* convert the "date" field to POSIX date type
* keep the data subset without missing values in **stepsReported** variable

```{r}
data <- read.csv(unz("activity.zip", 'activity.csv'))
data$date <- as.POSIXct(data$date)
stepsReported <- data[!is.na(data$steps),]
```

## What is mean total number of steps taken per day?

* calculate the total number of steps per day
* plot a histogram of steps per day 

```{r}
stepsByDate <- aggregate(stepsReported$steps, list(date = stepsReported$date), sum)
hist(stepsByDate$x, 
     col = "blue",
     main = "Total number of steps taken each day",
     xlab = "steps")
```

* calculate mean of total steps per day
```{r}
mean(stepsByDate$x)
```

* calculate median of total steps per day
```{r}
median(stepsByDate$x)
```


## What is the average daily activity pattern?

* calculate average number of steps per 5-min interval over all days
* plot average number of steps per interval 

``` {r}
avgStepsByInterval <- aggregate(stepsReported$steps, list(interval = stepsReported$interval), FUN = mean)
plot(avgStepsByInterval$interval, avgStepsByInterval$x, 
     type = "l", col='blue', 
     main = "Average daily activity pattern", 
     xlab = "5 minute interval", 
     ylab = "number of steps")
```

* find the interval containing the max number of steps
``` {r}
which.max(avgStepsByInterval$x)
```


## Imputing missing values

* calculate total number of missing values

``` {r}
sum(is.na(data$steps))
```

* create a new dataset 
* fill missing values by using average for the same 5 minute interval

``` {r}
newData <- data
for (i in  which(is.na(newData$steps))) {
    interval <- newData$interval[i]
    newData$steps[i] <- avgStepsByInterval[avgStepsByInterval$interval == interval,]$x
}
```

* calculate the total number of steps per day after imputting missing data 
* plot a histogram of steps per day 

```{r}
newStepsByDate <- aggregate(newData$steps, list(date = newData$date), sum)
hist(newStepsByDate$x, 
     col = "blue",
     main = "Total number of steps taken each day (after filling missing values)",
     xlab = "steps")
```

* calculate mean of total steps per day
```{r}
mean(newStepsByDate$x)
```

* calculate median of total steps per day
```{r}
median(newStepsByDate$x)
```


## Are there differences in activity patterns between weekdays and weekends?
